---
title: "Chicago Crime Data Analysis"
author: "Pankaj, Sakshi and Prafful"
date: "12/19/2021"
---

##### Importing Data and packages
```{r message=FALSE}
# add packages you need for this assignment
library("xlsx")
library(tidyverse)  # includes tibbles, ggplot2, dyplr, and more. 
```

> Importing required files from the path

```{r}
ccdata<- read.csv("~/Desktop/CC/ChicagoCrimeData20162021.csv") 
crime<- read.csv("~/Desktop/CC/crime1.csv", header=TRUE)
yearcc <- read.xlsx("~/Desktop/CC/yeardata16-21.xlsx", 1, header=TRUE)
```


```{r}
plot1 <-ggplot(yearcc, aes(x=Date)) + 
        geom_line(aes(y = Year, color = Date)) 
print(plot1 +ggtitle("CRIME IN CHICAGO FROM 2016 TO 2021")+ labs(y="No Of Incidents", x = "Timeline"))
```

> The above graph illustrates number of incidents occurred each year from 2016 to 2021. It clearly depicts that the highest number of incidents happened in 2018 and every year follows the same pattern. In the begning of every year there is a significant decline in number of incidents, as you can see in the Jan-Feb of 2016 the numbers are least and it gradually increases from March to August and achieve its peak during September to November and starts declining again from December. Suprisingly this pattern is identical for every year, we might notice overall decline in cases from 2020 but the pattern is exactly same.


```{r, fig.align='center'}
CrimeType <- table(ccdata$Primary.Type, ccdata$Year) ## Making a table between Primary crime type and year in which it occurred.
typecc <- data.frame(CrimeType) # making it a data frame

plot2 <- ggplot(data = typecc) +
  geom_point(mapping = aes(x = Var2, y = Freq,colour = Var1))
print(plot2 +ggtitle("TYPES OF CRIME IN CHICAGO FROM 2016 TO 2021")+ labs(y="Count of Incident", x = "Timeline", color = "Types"))
```

> The above graph is to determine the type of incident that happend during the time period of 2016 - 2021. From the graph we can conclude that "Theft" as a crime type has highest occurence followed by "Battery","Criminal Damage" "Deceptive Practice" adn so on. While these were some top crimes commited there are some types which were less than 100. To better understand the figures following is the tabular view of the types of crime segregated by Year, sorted alphabatically.


```{r}
CrimeType
```
```{r}
tt1 <- table(ccdata$YEAR,ccdata$REGION)
tt1
head(ccdata$Block)
```

```{r}
tt1 <- table(crime$YEAR,crime$REGION)

regioncc <- data.frame(tt1)

names(regioncc)[1] <- "Year"
names(regioncc)[2] <- "Region"
names(regioncc)[3] <- "Count"

regioncc$Region <- as.character(regioncc$Region)
regioncc$Year <- as.character.Date(regioncc$Year)

plot4 <- ggplot(data = regioncc) +
  geom_point(mapping = aes(x = Year, y = Count,colour = Region))
print(plot4 +ggtitle("No of Incidents in 4 Regions")+ labs(y="Count of Incident", x = "Timeline"))

```

> The above graph is between number of incidents recorded and region in which it took place. The scatter plot clearly illustrates that highest number of incidents were in the South region followed by West, North and East. However` we can notice decline in the recorded cases as we progress in the year but the sequence of region is identical every year.

```{r}
tt2 <- table(ccdata$Year,ccdata$Location.Description)
locationcc <- data.frame(tt2)

names(locationcc)[1] <- "Year"
names(locationcc)[2] <- "Location"
names(locationcc)[3] <- "Count"

locationcc$Year <- as.character.Date(locationcc$Year)
locationcc$Location <- as.character(locationcc$Location)
locationcc$Count <- as.numeric(locationcc$Count)

loc <- subset(locationcc,locationcc$Location == "APARTMENT" | locationcc$Location == "RESIDENCE" | locationcc$Location == "STREET" | locationcc$Location == "SIDEWALK"  )
loc

```





```{r}
data2016<-subset(crime,crime$YEAR == 2016)
data2017<-subset(crime,crime$YEAR == 2017)
data2018<-subset(crime,crime$YEAR == 2018)
data2019<-subset(crime,crime$YEAR == 2019)
data2020<-subset(crime,crime$YEAR == 2020)
data2021<-subset(crime,crime$YEAR == 2021)

```



```{r}

crime2016<-data.frame( table(data2016$PRIMARYTYPE))
names(crime2016)[1]<-'CrimeType'


crime2017<-data.frame( table(data2017$PRIMARYTYPE))
names(crime2017)[1]<-'CrimeType'


crime2018<-data.frame( table(data2018$PRIMARYTYPE))
names(crime2018)[1]<-'CrimeType'

crime2019<-data.frame( table(data2019$PRIMARYTYPE))
names(crime2019)[1]<-'CrimeType'


crime2020<-data.frame( table(data2020$PRIMARYTYPE))
names(crime2020)[1]<-'CrimeType'


crime2021<-data.frame( table(data2021$PRIMARYTYPE))
names(crime2021)[1]<-'CrimeType'


```


>HYPOTHESIS 1


>To check whether the crime type committed most in year 2016 is the same type of crime committed in year 2017 using Hypothesis Testing. Here we are applying prop-test through which we will get the p-value. And on the basis of p-value we can come to a conclusion. We are using prop-test because to get accurate data with respect to the total number of crime.

\begin{align*}
H_0: \mu1 & = \mu2, \\
H_1: \mu1 & \ne \mu2. 
\end{align*}


> H_0 is Null Hypothesis and H_1 is Alternative Hypothesis.

```{r}
#Hypothesis 1

crime2016$CrimeType[which.max(crime2016$Freq)] # Primary type in 2016

theft2016<-subset(crime2016$Freq,crime2016$CrimeType == 'THEFT')
theft2016

theft2017<-subset(crime2017$Freq,crime2017$CrimeType == 'THEFT')
theft2017


prop.test(x = c(theft2016,theft2017), n = c(nrow(data2016),nrow(data2017)), alternative = "greater", conf.level = 0.95)

```

>For the year 2016 THEFT is the most committed crime. The p-value of this hypothesis is greater than alpha i.e., 0.05. So we can accept the null hypothesis H_0 and agree that crime type committed most in 2016 is the same crime type committed in 2017.

>HYPOTHESIS 2 

>To check whether the crime committed most in a Region in year 2016 is the same Region for the year 2017 using Hypothesis Testing. Here we are applying prop-test through which we will get the p-value. And on the basis of p-value we can come to a conclusion.

\begin{align*}
H_0: \mu1 & = \mu2, \\
H_1: \mu1 & \ne \mu2. 
\end{align*}

> H_0 is Null Hypothesis and H_1 is Alternative Hypothesis.

```{r}

Region2016<-data.frame(table(data2016$REGION))
names(Region2016)[1]<-'Region'
Region2016

Region2017<-data.frame(table(data2017$REGION))
names(Region2017)[1]<-'Region'
Region2017

Region2016$Region[which.max(Region2016$Freq)] # region in 2016


South2016<-subset(Region2016$Freq,Region2016$Region == 'South')
South2016

South2017<-subset(Region2017$Freq,Region2017$Region == 'South')
South2017


prop.test(x = c(South2016,South2017), n = c(nrow(data2016),nrow(data2017)), alternative = "greater", conf.level = 0.95)

```


>The p-value of this hypothesis is smaller than alpha i.e., 0.05 but mu1 is equal to mu2.So, we can accept the null hypothesis H_0 and agree that the Region in which crime committed most in 2016 is the same Region for year 2017 where crime were committed most i.e., Region South.



>>HYPOTHESIS 3

>To check whether the domestic crime rate in 2020 increased or not in comparison to 2019 we are using Hypothesis Testing. As per my assumption domestic cases should increase because during lockdown people mostly stayed  at home.  Here we are applying prop-test through which we will get the p-value. And on the basis of p-value we can come to a conclusion.

\begin{align*}
H_0: \mu1 & = \mu2, \\
H_1: \mu1 & \ne \mu2. 
\end{align*}

> H_0 is Null Hypothesis and H_1 is Alternative Hypothesis.

```{r}
#Hypothesis 3
Domestic2019<-subset(crime,crime$YEAR == 2019 & crime$DOMESTIC_01 == 1)
nrow(Domestic2019)

Domestic2020<-subset(crime,crime$YEAR == 2020 & crime$DOMESTIC_01 == 1)
nrow(Domestic2020)


prop.test(x = c(nrow(Domestic2019),nrow(Domestic2020)), n = c(nrow(data2019),nrow(data2020)), alternative = "greater", conf.level = 0.95)

```


>The p-value of this hypothesis is greater than alpha i.e., 0.05 so, we can accept the NULL hypothesis H_0 and agree that the Domestic Violence cases increased in year 2020 in comparison to the year of 2019. Additonally I am using line graph to visualize the pattern of domestic cases from 2016 - 2021.

```{r}
domesticcc <- read.xlsx("~/Desktop/CC/Domesticcc.xlsx", 1, header=TRUE)

plot3 <-ggplot(domesticcc, aes(x=Year)) + 
        geom_line(aes(y = TRUE., color = "True"))
print(plot3 +ggtitle("DOMESTIC CRIME IN CHICAGO FROM 2016 TO 2021")+ labs(y="Count of Incidents", x = "Timeline"))
```

> From the graph we can see that the cases around 43000 in 2016 and went upto 44000 in 2018 but we saw a sudden decline in cases in 2019 achieving its all time low of 40000. However, the number started increasing from the begning of 2020 and incrased by 2000 in 2021 which directly relate to lockdown amid pandemic.


>HYPOTHESIS 4

>To check whether the number of arrest in year 2017  increased or decreased in comparsion to  year 2016. We will check this by using Hypothesis Testing. Here we are applying prop-test through which we will get the p-value. And on the basis of p-value we can come to a conclusion.

\begin{align*}
H_0: \mu1 & = \mu2, \\
H_1: \mu1 & \ne \mu2. 
\end{align*}

> H_0 is Null Hypothesis and H_1 is Alternative Hypothesis.

```{r}
#Hypothesis 4
Arrest2016<-subset(crime,crime$YEAR == 2016 & crime$ARREST_01 == 1)
nrow(Arrest2016)

Arrest2017<-subset(crime,crime$YEAR == 2017 & crime$ARREST_01 == 1)
nrow(Arrest2017)


prop.test(x = c(nrow(Arrest2016),nrow(Arrest2017)), n = c(nrow(data2016),nrow(data2017)), alternative = "greater", conf.level = 0.95)

```

>The p-value of this hypothesis is smaller than alpha i.e., 0.05 and the mu1 is also not equal to mu2. So, the proportion is also not equal. So we can accept the Alternate hypothesis H_1 and agree that the less criminal got arrested in year 2017 incomparison to the year 2016.




```{r}

LocationDesc2016<-data.frame(sort(table(data2016$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2016)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2016<- cbind(LocationDesc2016,Year=c(2016))


LocationDesc2017<-data.frame(sort(table(data2017$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2017)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2017<- cbind(LocationDesc2017,Year=c(2017))



LocationDesc2018<-data.frame(sort(table(data2018$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2018)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2018<- cbind(LocationDesc2018,Year=c(2018))



LocationDesc2019<-data.frame(sort(table(data2019$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2019)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2019<- cbind(LocationDesc2019,Year=c(2019))


LocationDesc2020<-data.frame(sort(table(data2020$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2020)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2020<- cbind(LocationDesc2020,Year=c(2020))


LocationDesc2021<-data.frame(sort(table(data2021$LOCATIONDESCRIPTION),decreasing = TRUE))
names(LocationDesc2021)[1]<- 'LOCATIONDESCRIPTION'
LocationDesc2021<- cbind(LocationDesc2021,Year=c(2021))


Top5allyears<- merge(merge(merge(merge(merge(LocationDesc2016[1:5,1:3],LocationDesc2017[1:5,1:3],all = TRUE,sort = FALSE),LocationDesc2018[1:5,1:3],all = TRUE,sort = FALSE),LocationDesc2019[1:5,1:3],all = TRUE,sort = FALSE),LocationDesc2020[1:5,1:3],all = TRUE,sort = FALSE),LocationDesc2021[1:5,1:3],all = TRUE, sort = FALSE)

Top5allyears
```
>From the above result we can see the top 5 crime location for each year. From 2016 to 2019 top 5 crime location are same for all the 4 years i.e., Street, Residence, Apartment, Sidewalk, Other. For the year 2020 Other crime location is replaced by Small Retail Shop crime location in the top 5 list and for year 2021 Small Retail Shop crime location is replaced by Parking Lot/ Garage crime location in top 5 crime location.

```{r}

plot6 <-ggplot(Top5allyears, aes(x=Year)) + 
        geom_line(aes(y = Freq, color = LOCATIONDESCRIPTION))
print(plot6 +ggtitle("Top 5 Crime location from 2016 TO 2021")+ labs(y="Count of Incidents", x = "Timeline"))
```


> The above graph illustrate top 5 loation in the city of Chicago where crimes took place. From the year 2016 'Street' is dominating the graph with 60k cases followed by 'Residence', 'Apartment'and 'Sidewalk'. We can also see the decline in the cases from 2019 and increase in 'Apartment' cases from 2020.


> This project idetifies the pattern in crimes commited by identifying the month it is commited in, the location where it was commited and the region where it was committed.
